---
title: "Lab 01 - Hello R"
author: "Anaelle Gackiere"
date: "01/14/2026"
output: 
  html_document: 
    fig_width: 6
    theme: cosmo
    highlight: espresso
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(datasauRus)
data("datasaurus_dozen")
```

## Exercises

### Exercise 1
data("datasaurus_dozen")

datasaurus_dozen %>%
  count(dataset) %>%
  print()
#> # A tibble: 13 × 2
#>    dataset        n
#>    <chr>      <int>
#>  1 away         142
#>  2 bullseye     142
#>  3 circle       142
#>  4 dino         142
#>  5 dots         142
#>  6 h_lines      142
#>  7 high_lines   142
#>  8 slant_down   142
#>  9 slant_up     142
#> 10 star         142
#> 11 v_lines      142
#> 12 wide_lines   142
#> 13 x_shape      142

### Exercise 2

The answers for this exercise are already provided to help guide you. Clean up the narrative to reflect what you want to submit. Let’s break it into steps:

First, let's plot the data in the dino dataset:

```{r plot-dino, fig.height=3, fig.width=6}
dino_data <- datasaurus_dozen %>%
  filter(dataset == "dino")

ggplot(data = dino_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

Next, let's calculate the correlation between `x` and `y` in this dataset:

```{r cor-dino}
dino_data %>%
  summarize(r = cor(x, y))
```
Seems like there is no significant relationship between x and y! 
And yet...the visuals tell a different story :-)


### Exercise 3

```{r plot_star}
star_data <- datasaurus_dozen %>%
  filter(dataset == "star")

ggplot(data = star_data, mapping = aes(x = x, y = y)) +
  geom_point()

```

Given our graph, it doesn't seem like we'll have a significant Pearson r.
But why not double check!

```{r cor_star}
star_data %>%
  summarize(r = cor(x, y))
```
There may be a non-linear relationship here; the points seem to diverge in a meaningful
way (away from the center), but the pearson r cannot specify this (as it mostly checks for linear relationships). 


### Exercise 4

As you'll see below, the circle data shares a similar correlation to that of the dino's, despite the fact that they present very differently visually. 

```{r plot_circle}
circle_data <- datasaurus_dozen %>%
  filter(dataset == "circle")

ggplot(data = circle_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

```{r cor_circle}
circle_data %>%
  summarize(r = cor(x, y))
```

### Exercise 5

As you'll see below, these graphs are wonderful to look at. Though they may seem so different from one another, they share very similar correlation coefficients. How fun!

```{r plot_all}
ggplot(datasaurus_dozen, aes(x = x, y = y, color = dataset)) +
  geom_point() +
  facet_wrap(~dataset, ncol = 3) +
  theme(legend.position = "none")
```


```{r cor_summary}
datasaurus_dozen %>%
  group_by(dataset) %>%
  summarize(r = cor(x, y)) %>%
  print(13)

```



